---
title: "Thesis"
author: "Grace Manninen"
date: "2025-11-28"
output: html_document
---
Study Design

(Two-tailed)
H₀: There is no significant difference in species ratings between the music condition and the non-music condition.

H₁: There is a significant difference in species ratings between the music condition and the non-music condition.

2x2 Factor Design - Within-Subjects

Independent Variable:
1: Music (Two-Levels)
- music
- no music
2: Species (Two-Levels)
- favoured
- unfavoured

Dependent Variable
Likert-Scale Question: “I would want to care for or protect this animal.”
1 = Strongly disagree
2 = Disagree
3 = Somewhat disagree
4 = Neither agree nor disagree
5 = Somewhat agree
6 = Agree
7 = Strongly agree

POWER ANALYSIS
```{r}
install.packages("pwr")
library(pwr)
```

alpha = 0.05 (significance threshold)
beta = 0.80 (power above 80%)
d = 0.5 (medium effect size)

```{r}
alpha <- 0.05
power_target <- 0.8
d_main <- 0.5
d_interaction <- 0.2
```

Main Effect of Music
```{r}
power_music <- pwr.t.test(d = d_main,
                          power = power_target,
                          sig.level = alpha,
                          type = "paired",
                          alternative = "greater")

cat("Main Effect of Music:\n")
print(power_music)
cat("\nNumber of participants needed for Music effect:", ceiling(power_music$n), "\n\n")

```
Number of participants needed for Music effect: 27 - 34 + 

Main Effect of Species
```{r}
power_species <- pwr.t.test(d = d_main,
                          power = power_target,
                          sig.level = alpha,
                          type = "paired",
                          alternative = "greater")

cat("Main Effect of Species:\n")
print(power_music)
cat("\nNumber of participants needed for species effect:", ceiling(power_music$n), "\n\n")
```
Number of participants needed for species effect: 27 - 34 +

Interaction Effect
```{r}
power_interaction <- pwr.t.test(d = d_interaction,
                                power = power_target,
                                sig.level = alpha,
                                type = "paired",
                                alternative = "two.sided")

cat("Interaction Effect (approximate):\n")
print(power_interaction)
cat("\nNumber of participants needed for Interaction effect:", ceiling(power_interaction$n), "\n\n")
```
Number of participants needed for Interaction effect: 199

Study should have around 40 participants (34 min)

exclude people who gave same likert scale response, or inadequate responses

