---
title: "analysis code"
author: "Grace Manninen"
date: "2025-03-24"
output: html_document
---
loading in packages/libraries
```{r}
library(dplyr)
library(stringr)
library(jsonlite)
```


loading in data 
```{r}
#LOAD IN DATA
# Block 1 data: data 1: This data looks at charismatic - likert - scary - likert 
# Block 2 data: data 2: This data looks at fixation cross - scary - likert

data1 <- read.csv("graces_simulateddata_analysis.csv")
data2 <- read.csv("fixation-scary.csv")
```

data 1
```{r}
#inspect the data 
class(data1)
#"data.frame"

#getting image rows and extacting file names
image_rows1 <- grep("<img src=", data1$stimulus)
print(image_rows1) 
#5  7  9 11 13 15 17 19 21 23 25 27 29 31 33 35

image_filenames1 <- gsub('.*<img src=\\"|\\\".*', '', data1$stimulus[image_rows1])
print(image_filenames1) 
# [1] "charismatic/wolf.jpg"        "scary/wasp.jpg"             
# [3] "charismatic/zebra.jpg"       "scary/snake.jpg"            
# [5] "charismatic/elephant.jpg"    "scary/cockroach.jpg"        
# [7] "charismatic/cheetah.jpeg"    "scary/bull.jpg"             
# [9] "charismatic/lion.jpg"        "scary/roundworm.png"        
# [11] "charismatic/giant_panda.jpg" "scary/maggot.jpg"           
# [13] "charismatic/gorilla.jpg"     "scary/tapeworm.jpg"         
# [15] "charismatic/tiger.jpg"       "scary/spider.jpg"      

#because likert responses follow image file... to identify likert responses

likert_rows1 <- image_rows1 + 1
print(likert_rows1)
#6  8 10 12 14 16 18 20 22 24 26 28 30 32 34 36

funny <- str_extract(data1$response[likert_rows1], '"funny":([0-5])') %>%
  str_remove('"funny":') %>% as.integer()

motherly <- str_extract(data1$response[likert_rows1], '"motherly":([0-5])') %>%
  str_remove('"motherly":') %>% as.integer()

friendly <- str_extract(data1$response[likert_rows1], '"friendly":([0-5])') %>%
  str_remove('"friendly":') %>% as.integer()

cute <- str_extract(data1$response[likert_rows1], '"cute":([0-5])') %>%
  str_remove('"cute":') %>% as.integer()

protective <- str_extract(data1$response[likert_rows1], '"protective":([0-5])') %>%
  str_remove('"protective":') %>% as.integer()

help <- str_extract(data1$response[likert_rows1], '"help":([0-5])') %>%
  str_remove('"help":') %>% as.integer()

image_likert_df <- data.frame(
  Image = image_filenames1,
  Funny = funny,
  Motherly = motherly,
  Friendly = friendly,
  Cute = cute,
  Protective = protective,
  Help = help
)

print(image_likert_df)

scary_images <- image_likert_df[grepl("scary", image_likert_df$Image), ]
print(scary_images)

#now scary_images displays a table that shows the scary animals with their corresponding likert scale responses

```

data 2
```{r}
#view data
head(data2)
colnames(data2)

#clean data, take out fixation cross data
stimulus_rows <- grep("<div style=\"font-size:60px;\">\\+<\\/div>", data2$stimulus)
data_clean2 <- data2[-stimulus_rows, ]
head(data_clean2)

#extracting file names
image_rows2 <- grep("<img src=", data_clean2$stimulus)
image_filenames2 <- gsub('.*<img src=\\"|\\\".*', '', data_clean2$stimulus[image_rows2])
print(image_filenames2)

#because likert responses follow image file... to identify likert responses

likert_rows2 <- image_rows2 + 1
print(likert_rows2)

head(data_clean2$response[likert_rows2])

likert_responses2 <- lapply(data_clean2$response[likert_rows2], fromJSON)
##going to check some rows of likert response corresponding to the image

print(likert_responses2[[1]]) #roundworm - matches
print(likert_responses2[[2]]) #tapeworm - matches
print(likert_responses2[[3]]) #maggot - matches
#... looks good

data_combined2 <- data.frame(
  Image = image_filenames2,
  Funny = sapply(likert_responses2, function(x) x$funny),
  Motherly = sapply(likert_responses2, function(x) x$motherly),
  Friendly = sapply(likert_responses2, function(x) x$friendly),
  Cute = sapply(likert_responses2, function(x) x$cute),
  Protective = sapply(likert_responses2, function(x) x$protective),
  Help = sapply(likert_responses2, function(x) x$help),
  stringsAsFactors = FALSE
)

```

now the data is sorted and cleaned
check once more before analysis
```{r}
#data1 
data1table <- scary_images
print(data1table)

#data2
data2table <- data_combined2
print(data2table)
```

I have clean and organized table's of data that has the information I want to compare block 1 and block 2

To run the Mann-Whitney U test, participants data that has N/A values or unanswered questions will be discarded. Also partcipants data that has the same answer put in a row will be discarded. Enough sets of **complete (from the above criteria) data for each block (n=44) will need to be collected. 

Each data set, will go through the above (respective if they are block 1 or block 2), to get rid of unnecessary data collected, to prepare for the inferential tests. 

To prepare for the Mann-Whitney U test

HYPOTHESIS: 
H0 (null hypothesis): Block 1 and Block 2 will not have a significant difference, failing to reject the null hypothesis
H1 (alternative hypothesis): Block 1 and Block 2 will be significantly different, rejecting the null hypothesis

creating a new data frame
- to simplify it: it will include simulated data where there is going to be an assumed effect of serial dependence. There will be no missing data (no N/A). 
```{r}
datafinal <- data.frame(
  participantID = 1:88,
  condition = rep(c("Block1", "Block2"), each=44),
  funny = rep(0, each=88),
  cute = rep(0, each=88),
  friendly = rep(0, each = 88),
  help = rep(0, each=88),
  motherly = rep(0, each=88),
  protective = rep(0, each=88),
  overall=rep(0, each=88)
)
```

[reminder that this simulated data for block 1 and block 2 is participants average for what they scored for each likert scale question, for all of the stimuli (scary images)]

block 1 -> should have a range of 1:5 for the responses, since assuming an effect of serial dependence, charismatic stimuli will cause scary stimuli to be more liked

block 2 -> should have a smaller and lower range, 1:4, for the responses, since assuming there will not be an effect of serial dependence, since fixation cross stimuli (control) is being shown instead of charismatic animal stimuli

FUNNY -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$funny[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$funny[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


CUTE -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$cute[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, not seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$cute[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


FRIENDLY -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$friendly[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$friendly[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


HELP -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$help[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$help[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


MOTHERLY -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$motherly[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$motherly[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


PROTECTIVE -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$protective[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$protective[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```

overall scores for each participant
```{r}
datafinal$overall <- rowMeans(datafinal[,3:9])
```

Planned Analysis:
Mann-Whitney U test for overall 
... now I have a table (datafinal) with simulated data, that should show me that block 1 will have higher scores (likert scale question (6), all on a scale of (1-5)), as higher scores mean in general that the animal is in general rated higher (cuter, friendlier, etc).

```{r}
overall1 <- datafinal$overall[datafinal$condition=="Block1"]
overall2 <- datafinal$overall[datafinal$condition=="Block2"]
```

Check what the average is for block 1 vs block 2
```{r}
mean(overall1) #3.022727
mean(overall2) #2.386364

```

the results are what I would have expected, it makes sense that the overall score for block 1 is higher then block 2

but I want to test if this is a significant effect, so I will do a Mann-Whitey U test (Wilcoxon rank-sum test)

OVERALL

```{r}
wilcox.test(overall1, overall2)

#	Wilcoxon rank sum test with continuity correction
#data:  overall1 and overall2
#W = 1540, p-value = 1.656e-06
#alternative hypothesis: true location shift is not equal to 0
```

can check other effects too...

FUNNY

```{r}
funny1 <- datafinal$funny[datafinal$condition=="Block1"]
funny2 <- datafinal$funny[datafinal$condition=="Block2"]

wilcox.test(funny1, funny2)

#	Wilcoxon rank sum test with continuity correction
#data:  funny1 and funny2
#W = 1035, p-value = 0.5695
#alternative hypothesis: true location shift is not equal to 0
```

CUTE

```{r}
cute1 <- datafinal$cute[datafinal$condition=="Block1"]
cute2 <- datafinal$cute[datafinal$condition=="Block2"]

wilcox.test(cute1, cute2)
#Wilcoxon rank sum test with continuity correction
#data:  cute1 and cute2
#W = 1341, p-value = 0.001484
#alternative hypothesis: true location shift is not equal to 0
```

FRIENDLY

```{r}
friendly1 <- datafinal$friendly[datafinal$condition=="Block1"]
friendly2 <- datafinal$friendly[datafinal$condition=="Block2"]

wilcox.test(friendly1, friendly2)

#Wilcoxon rank sum test with continuity correction
#data:  friendly1 and friendly2
#W = 1319, p-value = 0.002697
#alternative hypothesis: true location shift is not equal to 0
```

HELP

```{r}
help1 <- datafinal$help[datafinal$condition=="Block1"]
help2 <- datafinal$help[datafinal$condition=="Block2"]

wilcox.test(help1, help2)

#	Wilcoxon rank sum test with continuity correction
#data:  help1 and help2
#W = 1155, p-value = 0.1094
#alternative hypothesis: true location shift is not equal to 0
```

MOTHERLY

```{r}
motherly1 <- datafinal$motherly[datafinal$condition=="Block1"]
motherly2 <- datafinal$motherly[datafinal$condition=="Block2"]

wilcox.test(motherly1, motherly2)

#	Wilcoxon rank sum test with continuity correction
#data:  motherly1 and motherly2
#W = 1171.5, p-value = 0.08221
#alternative hypothesis: true location shift is not equal to 0
```

PROTECTIVE

```{r}
protective1 <- datafinal$protective[datafinal$condition=="Block1"]
protective2 <- datafinal$protective[datafinal$condition=="Block2"]

wilcox.test(protective1, protective2)

#	Wilcoxon rank sum test with continuity correction
#data:  protective1 and protective2
#W = 1265.5, p-value = 0.01101
#alternative hypothesis: true location shift is not equal to 0
```




