---
title: "analysis code"
author: "Grace Manninen"
date: "2025-03-24"
output: html_document
---
loading in packages/libraries
```{r}
library(dplyr)
library(stringr)
library(jsonlite)
library(ggplot2)
```

loading in data 
```{r}
#LOAD IN DATA
# Block 1 data: data 1: This data looks at charismatic - likert - scary - likert 
# Block 2 data: data 2: This data looks at fixation cross - scary - likert

data1 <- read.csv("graces_simulateddata_analysis.csv")
data2 <- read.csv("fixation-scary.csv")

#setting seed 
set.seed(20)
```

data 1
```{r}
#inspect the data 
class(data1)
#"data.frame"

#getting image rows and extacting file names
image_rows1 <- grep("<img src=", data1$stimulus)
print(image_rows1) 
#5  7  9 11 13 15 17 19 21 23 25 27 29 31 33 35

image_filenames1 <- gsub('.*<img src=\\"|\\\".*', '', data1$stimulus[image_rows1])
print(image_filenames1) 
# [1] "charismatic/wolf.jpg"        "scary/wasp.jpg"             
# [3] "charismatic/zebra.jpg"       "scary/snake.jpg"            
# [5] "charismatic/elephant.jpg"    "scary/cockroach.jpg"        
# [7] "charismatic/cheetah.jpeg"    "scary/bull.jpg"             
# [9] "charismatic/lion.jpg"        "scary/roundworm.png"        
# [11] "charismatic/giant_panda.jpg" "scary/maggot.jpg"           
# [13] "charismatic/gorilla.jpg"     "scary/tapeworm.jpg"         
# [15] "charismatic/tiger.jpg"       "scary/spider.jpg"      

#because likert responses follow image file... to identify likert responses

likert_rows1 <- image_rows1 + 1
print(likert_rows1)
#6  8 10 12 14 16 18 20 22 24 26 28 30 32 34 36

funny <- str_extract(data1$response[likert_rows1], '"funny":([0-5])') %>%
  str_remove('"funny":') %>% as.integer()

motherly <- str_extract(data1$response[likert_rows1], '"motherly":([0-5])') %>%
  str_remove('"motherly":') %>% as.integer()

friendly <- str_extract(data1$response[likert_rows1], '"friendly":([0-5])') %>%
  str_remove('"friendly":') %>% as.integer()

cute <- str_extract(data1$response[likert_rows1], '"cute":([0-5])') %>%
  str_remove('"cute":') %>% as.integer()

protective <- str_extract(data1$response[likert_rows1], '"protective":([0-5])') %>%
  str_remove('"protective":') %>% as.integer()

help <- str_extract(data1$response[likert_rows1], '"help":([0-5])') %>%
  str_remove('"help":') %>% as.integer()

image_likert_df <- data.frame(
  Image = image_filenames1,
  Funny = funny,
  Motherly = motherly,
  Friendly = friendly,
  Cute = cute,
  Protective = protective,
  Help = help
)

print(image_likert_df)

scary_images <- image_likert_df[grepl("scary", image_likert_df$Image), ]
print(scary_images)

#now scary_images displays a table that shows the scary animals with their corresponding likert scale responses

```

data 2
```{r}
#view data
head(data2)
colnames(data2)

#clean data, take out fixation cross data
stimulus_rows <- grep("<div style=\"font-size:60px;\">\\+<\\/div>", data2$stimulus)
data_clean2 <- data2[-stimulus_rows, ]
head(data_clean2)

#extracting file names
image_rows2 <- grep("<img src=", data_clean2$stimulus)
image_filenames2 <- gsub('.*<img src=\\"|\\\".*', '', data_clean2$stimulus[image_rows2])
print(image_filenames2)

#because likert responses follow image file... to identify likert responses

likert_rows2 <- image_rows2 + 1
print(likert_rows2)

head(data_clean2$response[likert_rows2])

likert_responses2 <- lapply(data_clean2$response[likert_rows2], fromJSON)
##going to check some rows of likert response corresponding to the image

print(likert_responses2[[1]]) #roundworm - matches
print(likert_responses2[[2]]) #tapeworm - matches
print(likert_responses2[[3]]) #maggot - matches
#... looks good

data_combined2 <- data.frame(
  Image = image_filenames2,
  Funny = sapply(likert_responses2, function(x) x$funny),
  Motherly = sapply(likert_responses2, function(x) x$motherly),
  Friendly = sapply(likert_responses2, function(x) x$friendly),
  Cute = sapply(likert_responses2, function(x) x$cute),
  Protective = sapply(likert_responses2, function(x) x$protective),
  Help = sapply(likert_responses2, function(x) x$help),
  stringsAsFactors = FALSE
)

```

now the data is sorted and cleaned
check once more before analysis
```{r}
#data1 
data1table <- scary_images
print(data1table)

#data2
data2table <- data_combined2
print(data2table)
```

I have clean and organized table's of data that has the information I want to compare block 1 and block 2

To run the Mann-Whitney U test, participants data that has N/A values or unanswered questions will be discarded. Also partcipants data that has the same answer put in a row will be discarded. Enough sets of **complete (from the above criteria) data for each block (n=44) will need to be collected. 

Each data set, will go through the above (respective if they are block 1 or block 2), to get rid of unnecessary data collected, to prepare for the inferential tests. 

To prepare for the Mann-Whitney U test

HYPOTHESIS: 
H0 (null hypothesis): Block 1 and Block 2 will not have a significant difference, failing to reject the null hypothesis
H1 (alternative hypothesis): Block 1 and Block 2 will be significantly different, rejecting the null hypothesis

creating a new data frame
- to simplify it: it will include simulated data where there is going to be an assumed effect of serial dependence. There will be no missing data (no N/A). 
```{r}
datafinal <- data.frame(
  participantID = 1:88,
  condition = rep(c("Block1", "Block2"), each=44),
  funny = rep(0, each=88),
  cute = rep(0, each=88),
  friendly = rep(0, each = 88),
  help = rep(0, each=88),
  motherly = rep(0, each=88),
  protective = rep(0, each=88),
  overall=rep(0, each=88)
)
```

[reminder that this simulated data for block 1 and block 2 is participants average for what they scored for each likert scale question, for all of the stimuli (scary images)]

block 1 -> should have a range of 1:5 for the responses, since assuming an effect of serial dependence, charismatic stimuli will cause scary stimuli to be more liked

block 2 -> should have a smaller and lower range, 1:4, for the responses, since assuming there will not be an effect of serial dependence, since fixation cross stimuli (control) is being shown instead of charismatic animal stimuli

FUNNY -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$funny[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$funny[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


CUTE -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$cute[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, not seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$cute[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


FRIENDLY -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$friendly[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$friendly[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


HELP -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$help[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$help[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


MOTHERLY -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$motherly[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$motherly[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```


PROTECTIVE -> making simulated data to show my expected effect of serial dependence
```{r}

#SIMULATE FUNNY DATA
datafinal$protective[datafinal$condition=="Block1"] <- sample(1:5, 44, replace=TRUE)

#assuming bc of block 2, nnot seeing charismatic -> should have a lower average score (no serial dependence), only 1:4

datafinal$protective[datafinal$condition=="Block2"] <- sample(1:4, 44, replace=TRUE)

```

overall scores for each participant
```{r}
datafinal$overall <- rowMeans(datafinal[,3:9])
```

Planned Analysis:
Mann-Whitney U test for overall 
... now I have a table (datafinal) with simulated data, that should show me that block 1 will have higher scores (likert scale question (6), all on a scale of (1-5)), as higher scores mean in general that the animal is in general rated higher (cuter, friendlier, etc).

```{r}
overall1 <- datafinal$overall[datafinal$condition=="Block1"]
overall2 <- datafinal$overall[datafinal$condition=="Block2"]
```

Check what the average is for block 1 vs block 2
```{r}
mean(overall1) #3.022727
mean(overall2) #2.386364

```

the results are what I would have expected, it makes sense that the overall score for block 1 is higher then block 2

but I want to test if this is a significant effect, so I will do a Mann-Whitey U test (Wilcoxon rank-sum test)

OVERALL

```{r}
wilcox.test(overall1, overall2)
```
The difference between the overall scores show significant effects (W = 1533.5, p-value = 2.183e-06
)

can check other effects too...

FUNNY

```{r}
funny1 <- datafinal$funny[datafinal$condition=="Block1"]
funny2 <- datafinal$funny[datafinal$condition=="Block2"]

wilcox.test(funny1, funny2)
```
The difference between the funny scores show non-significant effects (W = 1078.5, p-value = 0.3481
)

CUTE

```{r}
cute1 <- datafinal$cute[datafinal$condition=="Block1"]
cute2 <- datafinal$cute[datafinal$condition=="Block2"]

wilcox.test(cute1, cute2)
```
The difference between the cute scores show non-significant effects (W = 1181, p-value = 0.06959
)

FRIENDLY

```{r}
friendly1 <- datafinal$friendly[datafinal$condition=="Block1"]
friendly2 <- datafinal$friendly[datafinal$condition=="Block2"]

wilcox.test(friendly1, friendly2)
```
The difference between the friendly scores show significant effects (W = 1243.5, p-value = 0.01899
)

HELP

```{r}
help1 <- datafinal$help[datafinal$condition=="Block1"]
help2 <- datafinal$help[datafinal$condition=="Block2"]

wilcox.test(help1, help2)
```
The difference between the help scores show significant effects (W = 1288.5, p-value = 0.006259
)

MOTHERLY

```{r}
motherly1 <- datafinal$motherly[datafinal$condition=="Block1"]
motherly2 <- datafinal$motherly[datafinal$condition=="Block2"]

wilcox.test(motherly1, motherly2)
```
The difference between the help scores show significant effects (W = 1277.5, p-value = 0.008229
)

PROTECTIVE

```{r}
protective1 <- datafinal$protective[datafinal$condition=="Block1"]
protective2 <- datafinal$protective[datafinal$condition=="Block2"]

wilcox.test(protective1, protective2)
```
The difference between the help scores show non-significant effects (W = 1189.5, p-value = 0.05834
)

-------------------------------------------------------------------------------
data visualization - for block 1 vs block 2
- making bar graphs for each likert scale question means to compare block 1 and block 2
- error bars is standard error (upper bound and lower bound)

OVERALL
```{r}
overall_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_overall = c(mean(overall1), mean(overall2))
)

ggplot(overall_means, aes(x = condition, y = mean_overall, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_overall - sd(overall1)/sqrt(length(overall1)), 
                    ymax = mean_overall + sd(overall2)/sqrt(length(overall2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Overall Ratings", x = "Condition", y = "Mean Overall Score")
```

FUNNY
```{r}
funny_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_funny = c(mean(funny1), mean(funny2))
)

ggplot(funny_means, aes(x = condition, y = mean_funny, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_funny - sd(funny1)/sqrt(length(funny1)), 
                    ymax = mean_funny + sd(funny2)/sqrt(length(funny2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Funny Ratings", x = "Condition", y = "Mean Funny Score")
```

CUTE
```{r}
cute_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_cute = c(mean(cute1), mean(cute2))
)

ggplot(cute_means, aes(x = condition, y = mean_cute, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_cute - sd(funny1)/sqrt(length(funny1)), 
                    ymax = mean_cute + sd(funny2)/sqrt(length(funny2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Cute Ratings", x = "Condition", y = "Mean Cute Score")
```

FRIENDLY
```{r}
friendly_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_friendly = c(mean(friendly1), mean(friendly2))
)

ggplot(friendly_means, aes(x = condition, y = mean_friendly, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_friendly - sd(friendly1)/sqrt(length(friendly1)), 
                    ymax = mean_friendly + sd(friendly2)/sqrt(length(friendly2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Friendly Ratings", x = "Condition", y = "Mean Friendly Score")
```
HELP
```{r}
help_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_help = c(mean(help1), mean(help2))
)

ggplot(help_means, aes(x = condition, y = mean_help, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_help - sd(help1)/sqrt(length(help1)), 
                    ymax = mean_help + sd(help2)/sqrt(length(help2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Help Ratings", x = "Condition", y = "Mean Help Score")
```
MOTHERLY
```{r}
motherly_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_motherly = c(mean(motherly1), mean(motherly2))
)

ggplot(motherly_means, aes(x = condition, y = mean_motherly, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_motherly - sd(motherly1)/sqrt(length(motherly1)), 
                    ymax = mean_motherly + sd(motherly2)/sqrt(length(motherly2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Motherly Ratings", x = "Condition", y = "Mean Motherly Score")
```

PROTECTIVE
```{r}
protective_means <- data.frame(
  condition = c("Block1", "Block2"),
  mean_protective = c(mean(protective1), mean(protective2))
)

ggplot(protective_means, aes(x = condition, y = mean_protective, fill = condition)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_errorbar(aes(ymin = mean_protective - sd(protective1)/sqrt(length(protective1)), 
                    ymax = mean_protective + sd(protective2)/sqrt(length(protective2))), width = 0.2) +
  theme_minimal() +
  labs(title = "Comparison of Protective Ratings", x = "Condition", y = "Mean Protective Score")
```
